---
title: "Reproducible Research: Peer Assessment 1"
output:
  html_document:
    keep_md: true
---



## Loading and preprocessing the data
```{r load_packages, echo=TRUE,  results="hide"}
require(dplyr)
require(ggplot2)
```

```{r read_data, echo=TRUE}
unzip('activity.zip')
activity <- read.csv('activity.csv')
```


## What is mean total number of steps taken per day?
```{r , echo=TRUE}
by_date <- group_by(activity, date)
part1 <- summarise(by_date, total = sum(steps, na.rm= TRUE) )
mean.steps <- round(mean(part1$total))
median.steps <- format(round(median(part1$total), digits = 0), scientific = FALSE)
```

```{r steps_by_day, echo=TRUE}
hist(part1$total, main="Number of steps by day", xlab="number of steps")
```


The mean of the total number of steps taken per day is on average of **`r mean.steps` steps**

The median of the total number of steps taken per day is on average of **`r median.steps` steps**

## What is the average daily activity pattern?
```{r daily_pattern_fig, echo=TRUE}
plot(activity$interval, activity$steps, type="l",xlab="interval", ylab="steps", main="Daily activity pattern")
```
```{r daily_pattern, echo=TRUE}
interval.group <- group_by(activity, interval)
interval.steps <- summarise(interval.group, mean(steps, na.rm = TRUE))

colnames(interval.steps) <- c("interval", "mean.steps")
most.steps <- interval.steps[which.max(interval.steps$mean.steps),]
```

The **`r most.steps$interval`** 5 minutes interval contains the most steps across the dataset.


## Imputing missing values
```{r missing_values, echo=TRUE}
missing.steps <- nrow(filter(activity, is.na(steps)))
```

Total number of missing values in the dataset: **`r missing.steps`**


```{r replace_na, echo=TRUE}
interval.group <- group_by(activity, interval)
interval.steps <- summarise(interval.group, mean(steps, na.rm = TRUE))

#replace na return the mean of the interval group for a given interval.
replace.na <- function(x)
{
  if(x > 2355 | x < 0 | (x %% 5 != 0)){
    return(NA)
  }else{
  return(interval.steps[which(interval.steps[,1]== x),2][[1]])
  }
}

#replacing every na with the mean of the corresponding 5 min group
#and create a modified dataframe (activity.mod).
activity.mod <- activity
for(i in 1:(dim(activity.mod)[1]) ){
  if(is.na(activity.mod[i,]$steps)){
    activity.mod[i,]$steps = replace.na(activity.mod[i,]$interval)
  }
}
#idem as part1
 by_date.mod <- group_by(activity.mod, date)
 part1.mod <- summarise(by_date.mod, total = sum(steps, na.rm= TRUE) )
mean.steps.mod <- round(mean(part1.mod$total), digits = 0)
median.steps.mod <- round(median(part1.mod$total), digits = 0)
```
```{r histogram_na_replaced, echo=TRUE}
hist(part1.mod$total, main="Number of steps by day (with NA replaced by the mean of the group", xlab="number of steps")
```

The mean of the total number of steps taken per day is on average of **`r format(mean.steps.mod, scientific=FALSE)` steps**

The median of the total number of steps taken per day is on average of **`r format(median.steps.mod, scientific=FALSE)` steps**


The substitution of missing data by their group mean doesn't modify  the mean or median value of the number of steps by days.
It render the distribution more normal.


## Are there differences in activity patterns between weekdays and weekends?

```{r weekly_patterns, echo=TRUE}

dow <- function(x)
{
  num.day <- as.POSIXlt(x)$wday
  if(num.day == 0 | num.day == 6)
  {
    return( "weekend")
  }else {
    return("weekday")
  }
}

activity.mod <- mutate(activity.mod, numday = as.POSIXlt(as.Date(date))$wday)
activity.mod <- mutate(activity.mod, doweek = ifelse(numday %in% c(0,6), "weekend","weekday"))

#There is probably a oneliner in R, but lacking of expertise, I separated the dataframe in two groups.
#merging them after calculating the mean steps by interval.
g1 <- subset(activity.mod, doweek == "weekend")
g2 <- subset(activity.mod, doweek == "weekday")

interval.g1 <- group_by(g1, interval)
interval.g2 <- group_by(g2, interval)
interval.steps.g1 <- summarise(interval.g1, mean(steps, na.rm = TRUE))
interval.steps.g2 <- summarise(interval.g2, mean(steps, na.rm = TRUE))
interval.steps.g1$doweek = "weekend"
interval.steps.g2$doweek = "weekday"
final.table <- rbind(interval.steps.g1, interval.steps.g2)
colnames(final.table) <- c("interval", "mean.steps", "doweek")
```
```{r weekday_weekend, echo=TRUE, fig.width=8, fig.height=6}
qplot(interval, mean.steps, data=final.table, facets = .~doweek, geom="line")
```

Eventually, there is a noticeable different pattern of activity between the weekdays, with a peak
of activity in the morning, probably going to work, then a quiet sequence, probably sitting at the desk !
The weekend is marked by a globally more active day, but without the big morning peak.
Anyway, very interesting data, pushing for having his/her own personal monitor!
