---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---



## Loading and preprocessing the data
```{r , echo=TRUE}
require(dplyr)
require(ggplot2)
unzip('activity.zip')
activity <- read.csv('activity.csv')
```


## What is mean total number of steps taken per day?
```{r , echo=TRUE}
 by_date <- group_by(activity, date)
 part1 <- summarise(by_date, total = sum(steps, na.rm= TRUE) )
mean.steps <- round(mean(part1$total))
median.steps <- round(median(part1$total), digits = 0)
hist(part1$total, main="Number of steps by day", xlab="number of steps")
```


The mean of the total number of steps taken per day is on average of **`r mean.steps` steps**

The median of the total number of steps taken per day is on average of **`r median.steps` steps**

## What is the average daily activity pattern?
```{r , echo=TRUE}
plot(activity$interval, activity$steps, type="l")
interval.group <- group_by(activity, interval)
interval.steps <- summarise(interval.group, mean(steps, na.rm = TRUE))

colnames(interval.steps) <- c("interval", "mean.steps")
most.steps <- interval.steps[which.max(interval.steps$mean.steps),]
```
TODO: put a R embedded  formula 

The 835 5 minutes interval contains the most steps across the dataset. 


## Imputing missing values
```{r , echo=TRUE}
missing.steps <- nrow(filter(activity, is.na(steps)))
```

Total number of missing values in the dataset: **`r missing.steps` **


```{r , echo=TRUE}
interval.group <- group_by(activity, interval)
interval.steps <- summarise(interval.group, mean(steps, na.rm = TRUE))
replace.na <- function(x)
{
  if(x > 2355 | x < 0 | (x %% 5 != 0)){
    return(NA)
  }else{
  return(interval.steps[which(interval.steps[,1]== x),2][[1]])
  }
}
#replacing every na with the mean of the corresponding 5 min group
activity.mod <- activity
for(i in 1:(dim(activity.mod)[1]) ){
  if(is.na(activity.mod[i,]$steps)){
    activity.mod[i,]$steps = replace.na(activity.mod[i,]$interval)
  }
}
#idem as part1
 by_date.mod <- group_by(activity.mod, date)
 part1.mod <- summarise(by_date.mod, total = sum(steps, na.rm= TRUE) )
mean.steps.mod <- round(mean(part1.mod$total), digits = 0)
median.steps.mod <- round(median(part1.mod$total), digits = 0)
hist(part1.mod$total, main="Number of steps by day (with NA replaced by the mean of the group", xlab="number of steps")
```

The mean of the total number of steps taken per day is on average of **`r mean.steps.mod` steps**

The median of the total number of steps taken per day is on average of **`r median.steps.mod` steps**


The substitution of missing data by their group mean doesn't modify  the mean or median value of the number of steps by days.
I render the distribution more normal.


## Are there differences in activity patterns between weekdays and weekends?

```{r , echo=TRUE}
dow <- function(x)
{
  num.day <- as.POSIXlt(x)$wday
  if(num.day == 0 | num.day == 6) 
  {
    return( "weekend")
  }else {
    return("weekday")
  }
}

activity.mod <- mutate(activity.mod, numday = as.POSIXlt(as.Date(date))$wday)
activity.mod <- mutate(activity.mod, doweek = ifelse(numday %in% c(0,6), "weekend","weekday"))

g1 <- subset(activity.mod, doweek == "weekend")
g2 <- subset(activity.mod, doweek == "weekday")

interval.g1 <- group_by(g1, interval)
interval.g2 <- group_by(g2, interval)
interval.steps.g1 <- summarise(interval.g1, mean(steps, na.rm = TRUE))
interval.steps.g2 <- summarise(interval.g2, mean(steps, na.rm = TRUE))
interval.steps.g1$doweek = "weekend"
interval.steps.g2$doweek = "weekday"
final.table <- rbind(interval.steps.g1, interval.steps.g2)
colnames(final.table) <- c("interval", "mean.steps", "doweek")
qplot(interval, mean.steps, data=final.table, facets = .~doweek, geom="line")
```

