---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r , echo=TRUE}
require(dplyr)
unzip('activity.zip')
activity <- read.csv('activity.csv')
```


## What is mean total number of steps taken per day?
```{r , echo=TRUE}
 by_date <- group_by(activity, date)
 part1 <- summarise(by_date, total = sum(steps, na.rm= TRUE) )
mean.steps <- round(mean(part1$total))
median.steps <- round(median(part1$total))
hist(part1$total, main="Number of steps by day", xlab="number of steps")
```
The mean of the total number of steps taken per day is on average of **`r mean.steps` steps**

The median of the total number of steps taken per day is on average of **`r median.steps` steps**

## What is the average daily activity pattern?
```{r , echo=TRUE}
plot(activity$interval, activity$steps, type="l")
```
The 800 5 minutes interval contains the most steps across the dataset. 


## Imputing missing values
```{r , echo=TRUE}
missing.steps <- nrow(filter(activity, is.na(steps)))
```
Total number of missing values in the dataset: **`r missing.steps` **


## Are there differences in activity patterns between weekdays and weekends?
